---
title: "3D Imaging"
output: html_document
date: "2024-08-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rgl)
library(pals)
```

## Including Plots

```{r}
setwd("C:/Users/ALIESCH/OneDrive - Environmental Protection Agency (EPA)/Desktop/Intermediate Rasters/Historic Land Use")

LandUse<-raster("Reprojected_LandUseRaster_igh.tif")

cls <- data.frame(id=1:8, cover=c("Cropland", "Forest", "Grassland", "Shrubland", "Urban", "Rock/Ice/Desert",  "Tundra", "Pasture"))

levels(LandUse) <- cls

cols <- c("yellow", "darkgreen", "green", "orange", "red", 'grey', 'white', 'lightgreen')

setwd("C:/Users/ALIESCH/OneDrive - Environmental Protection Agency (EPA)/Desktop/AgCensus 2022")

county<-vect("UScounties.shp")

states <- aggregate(county, by='STATE_NAME')

neb <- states %>% filter(STATE_NAME == 'Nebraska')

Neb_igh<-project(neb, LandUse)

Neb_sp <- as(Neb_igh, "Spatial")

Neb_LandUse<-raster::crop(x=LandUse, y=Neb_sp)

Neb_LU_m<-raster::as.matrix(Neb_LandUse)

col = cols[cut(Neb_LU_m, 8)]


Pasture_Mask <- raster:::.ifel(Neb_LandUse == 1, 1, 0)

Pasture_LU_m<-raster::as.matrix(Pasture_Mask)

pastCol<-c("white", "lightgreen")

past_col = pastCol[cut(Pasture_LU_m, 2)]

 # Neb_sp$id=as.character(1:nrow(Neb_sp))
open3d()

surface3d(1:1334, 1:2964, rep(0,length(past_col)), color = past_col, back = "lines", smooth=FALSE)
surface3d(1:1334, 1:2964, rep(1,length(col)), color = col, back = "lines", smooth=FALSE)

