---
title: "Data"
output: html_document
date: "2024-08-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyterra)
library(tidyverse)
library(terra)
library(exactextractr)
library(data.table)
library(htmlTable)
library(gt)
library(patchwork)
library(png)
library(cowplot)
```

This file was utilized to create a composite figure for the GCAM
```{r}
setwd("C:/Users/ALIESCH/OneDrive - Environmental Protection Agency (EPA)/Desktop/Intermediate Rasters/Historic Land Use")

LandUse<-rast("Reprojected_LandUseRaster_igh.tif")

cls <- data.frame(id=1:8, cover=c("Cropland", "Forest", "Grassland", "Shrubland", "Urban", "Rock/Ice/Desert",  "Tundra", "Pasture"))

levels(LandUse) <- cls

cols <- c("Cropland" = "yellow", "Forest" = "darkgreen", "Grassland" = "lavender", "Shrubland" = "orange", "Urban"='red', "Rock/Ice/Desert"='grey', "Tundra"='white', "Pasture"='lightgreen')

ggplot() +
  geom_spatraster(data=LandUse)+
  scale_fill_manual(values = cols, na.translate=FALSE)

setwd("C:/Users/ALIESCH/OneDrive - Environmental Protection Agency (EPA)/Desktop/Moved Fies/AgCensus 2022")

county<-vect("UScounties.shp")

states <- aggregate(county, by='STATE_NAME')

neb <- states %>% filter(STATE_NAME == 'Nebraska')

Neb_igh<-project(neb, LandUse)

Neb_LandUse<-crop(LandUse, Neb_igh)

neb_LU<-Neb_LandUse %>% drop_na(cover) 

nebLU<-ggplot() +
  geom_spatraster(data=neb_LU)+
  scale_fill_manual(values = cols, name="Land Use", na.translate=FALSE)

leg <- get_legend(nebLU)

ggsave(plot=leg, filename = "legend.png", height = 2.31, dpi = 300)

# Convert to a ggplot and print


nebLU.plt<-ggplot() +
  geom_spatraster(data=Neb_LandUse, show.legend = FALSE)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank()) +
  scale_fill_manual(values = cols)+
  geom_sf(data=Neb_igh, fill = 'NA', linewidth = 1.5)

ggsave(plot=nebLU.plt, filename = "nebLU_plt.png",width = 4.46, height = 2.31, dpi = 300)



Crop_Mask <- ifel(Neb_LandUse == 1, 1, NA)

clsPast <- data.frame(id=1, cover=c("Cropland"))

levels(Crop_Mask) <- clsPast

pastCol<- c(LUColors, "NA"=NA)

cropland.plt<-ggplot() +
  geom_spatraster(data=Crop_Mask, show.legend = FALSE) +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank()) +
  scale_fill_manual(values = pastCol)+
  geom_sf(data=Neb_igh, fill = 'NA', linewidth = 1.5)

ggsave(plot=cropland.plt, filename = "cropland_plt.png",width = 4.46, height = 2.31, dpi = 300)

setwd("C:/Users/ALIESCH/OneDrive - Environmental Protection Agency (EPA)/Documents/soilrastercomparison")

SG2020<-rast("rasters/SoilGrids2020_0-30.tif")

Neb_SOC<-crop(SG2020, Neb_igh)


SG2020_Crop<-mask(Neb_SOC, Crop_Mask)



soilCrop.plt<-ggplot() +
  geom_spatraster(data=SG2020_Crop) +
  labs(fill = "Mg C/ha") +
  theme(legend.position = c(0.12, 0.5)) +
  scale_fill_continuous(type = "viridis") +
  geom_sf(data=Neb_igh, fill = 'NA', linewidth = 1.5)

ggsave(plot=soilCrop.plt, filename = "soilCrop_plt.png",width = 4.46, height = 2.31, dpi = 300)
  


maskExtract<-mask(SG2020_Crop, Neb_igh)

Neb_sf<-sf::st_as_sf(Neb_igh)

mean<-terra::extract(
    maskExtract, Neb_igh, 
    fun="mean", na.rm=TRUE
)

exact_extract(maskExtract, Neb_sf, "mean", progress = FALSE)
exact_extract(maskExtract, Neb_sf, "quantile", quantiles=0.25, progress = FALSE)
exact_extract(maskExtract, Neb_sf, "quantile", quantiles=0.75, progress = FALSE)

hist<- terra::hist(maskExtract, breaks=50, maxcell=1500000)


dat <- data.frame(counts= hist$counts,breaks = hist$mids)

datN<-dat[-c(46:61),]
datC<-datN[-c(1:6),]

dat.C.plt<-ggplot(datC, aes(x = breaks, y = counts, fill =breaks)) + ## Note the new aes fill here
  geom_bar(stat = "identity",fill='black', alpha = 0.5)+
  xlab("Soil Carbon Bins")+ ylab("Count")+
  scale_x_continuous(breaks = seq(10,90,10),  ## without this you will get the same scale
                   labels = c(10, 20, 30, 40, 50, 60, 70, 80, 90))  +  ## as hist (question picture)
  geom_vline(xintercept = 57.15, color='red', size=1.5)

ggsave(plot=dat.C.plt, filename = "dat_C_plt.png",width = 4.46, height = 2.31, dpi = 300)


mx<-data.table(Statistic=c("Minimum","10th Percentile", "1st Quartile","Median", "Mean", "3rd Quartile", "90th Percentile", "Maximum"))



statTable <- mx %>%
  gt() %>%
  #tab_options(data_row.padding = px(1)) %>%
  opt_stylize(style = 1) %>%
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_body(
  )) %>%
  tab_style(
    style = cell_text(align = "center", weight = "bold", size="large"),
    locations = cells_column_labels()) %>%
 tab_style(
    style = cell_borders(
      sides = "all",
      color = "red",
      weight = px(1.5),
      style = "solid"
    ),
    locations = cells_body(rows = 6)
  )



gtsave(data=statTable, filename = "statTable.png")


statTable<-magick::image_read("statTable.png") 
statTable<-image_scale(statTable, "x225")


NEBLandUse<-magick::image_read("nebLU_plt.png")
NEBLandUse<-image_scale(NEBLandUse, "x225")

histSOC<-magick::image_read("dat_C_plt.png")
histSOC<-image_scale(histSOC, "x225")

soilCrop<-magick::image_read("soilCrop_plt.png")
soilCrop<-image_scale(soilCrop, "x225")

cropland<-magick::image_read("cropland_plt.png")
cropland<-image_scale(cropland, "x225")

legends<-magick::image_read("legend.png")
legends<-image_trim(legends)
legends<-image_scale(legends, "x200")

level1<-magick::image_append(c(NEBLandUse, legends, cropland))
level2<-magick::image_append(c(soilCrop, statTable, histSOC))

stacked<-c(level1, level2)

stacks<-magick::image_append(stacked, stack = TRUE)