---
title: "Historic Land Use"
output: html_document
date: "2023-01-11"
---

```{r}
library(terra)
library(tidyverse)
library(tidyterra)


```

Values we Use with GCAM
1: Cropland
2: Forest
3: Grassland
4: Shrubland
5: Urban Land
6: RID
7: Tundra
8: Pasture

Values used with this HYDE 3.2 Anthropocine project, and their equivalent classification:

"11 Urban" - Urban
"12 Dense settlements" - Urban

"21 Village, Rice" crop
"22 Village, Irrigated" crop 
"23 Village, Rainfed" crop
"24 Village, Pastoral" -Pasture

"31 Croplands, residential irrigated" - Crop
"32 Croplands, residential rainfed" - Crop
"33 Croplands, populated" - Crop 
"34 Croplands, pastoral" - Pasture

"41 Rangeland, residential" - Pasture
"42 Rangeland, populated" - grassland
"43 Rangeland, remote" - grassland

"51 Semi-natural woodlands, residential"  -forest
"52 Semi-natural woodlands, populated" - forest
"53 Semi-natural woodlands, remote" - forest
"54 Semi-natural treeless and barren lands" - grassland

"61 Wild, remote - woodlands"  forest
"62 Wild, remote - treeless & barren" RID
"63 Wild, remote - ice" - RID

"70 No definition"

```{r}
#Set the working directory
workingDir<-"C:/Users/aliesch/OneDrive - Environmental Protection Agency (EPA)/Desktop/Intermediate Rasters/Historic Land Use"
setwd(workingDir)

#Get all filenames with a .asc pattern
filenames <- list.files(workingDir, pattern="*.asc")


map <- function(index) {
  #Dial up the index valute
  year<-rast(filenames[index])
  RastUrban <- ifel(year == 11|year==12, 5, year)
  RastRID <- ifel(year == 62|year==63, 6, RastUrban)
  RastGrassland<- ifel(year == 54|year==42|year == 43, 3, RastRID)
  RastPasture<- ifel(year == 24|year==34|year==41, 8, RastGrassland)
  #RastShrub <- ifel(year == 43, 4, RastPasture)
  RastForest <- ifel(year == 61|year ==51|year == 52|year == 53, 2, RastPasture)
  RastCrop <- ifel(year== 21|year== 22|year== 23|year== 31|year== 32|year== 33, 1, RastForest)
  #Recolor the raster
  colors<-c("#fed976", "#00441b", "#c7e9c0", "#e41a1c", "#f0f0f0", "#9ecae1", "#41ab5d")
  #Reproject the Raster to igh
  #Load Tundra
  tundra<-rast("tundraProj.tif")
  RastTundra<-ifel(tundra == 1, 7, RastCrop)
  factTundra<-as.factor(RastTundra)
  years<-filenames[index]
  sub1<-str_sub(years, 10, -5)
  return(factTundra)
}

raster_output<-list()

index<-1:12

for (i in index){
  r<-map(i)
  raster_output[[i]]<-r
  }









```


```{r}

Rast1700<-rast("anthromes1700AD.asc")

plot(Rast1700)

LUR<-rast("Reprojected_LandUseRaster_igh.tif")

LURRast<-terra::project(Rast1700, LUR)



#Determine reclassification process

RastUrban <- ifel(Rast1700 == 11|Rast1700==12, 5, Rast1700)
plot(RastUrban)

RastRID <- ifel(Rast1700 == 62|Rast1700==63, 6, RastUrban)

RastGrassland<- ifel(Rast1700 == 54|Rast1700==42, 3, RastRID)

RastPasture<- ifel(Rast1700 == 24|Rast1700==34|Rast1700==41, 8, RastGrassland)

RastShrub <- ifel(Rast1700 == 43, 4, RastPasture)

RastForest <- ifel(Rast1700 == 61|Rast1700 ==51|Rast1700 == 52|Rast1700 == 53, 2, RastShrub)

RastCrop <- ifel(Rast1700== 21|Rast1700== 22|Rast1700== 23|Rast1700== 31|Rast1700== 32|Rast1700== 33, 1, RastForest)

LU1700<-RastCrop

colors<-c("#fed976", "#4daf4a", "#b2df8a", "#fdbf6f", "#e41a1c", "#f0f0f0", "#9ecae1", "#74c476")
plot(LU1700, col=colors)

tundra<-rast("tundraProj.tif")

RastTundra<-ifel(tundra == 1, 7, RastCrop)


Rast26<-ifel(LandUse == 26, NA)

breaks<-c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5)
stringer<-"anthromes1700AD.asc"
str_split(stringer, "s")
title<-str_sub(stringer, 10, -5)

plot(RastTundra)

fact<-as.factor(RastTundra)

LU1700<-terra::project(RastTundra, LUR)


```

```{r}





writeRaster(LU1700, "Recreations/LU1700.tif")

writeRaster(LU1800, "Recreations/LU1800.tif")

plot1800<-map(3)

plot1900<-map(5)

plot2000<-map(7)

```
