---
title: "GCAM Comparison"
author: "Mandy Liesch"
date: "5/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(terra)
library(tidyverse)
library(raster)
library(exactextractr)
library(sp)
library(rgdal)
library(sf)
library(tidyverse)

```
## GCAM Comparison vs Soil Grid 2017


```{r}
#Read in the GCAM Data
GCAM_SOC<-rast("GCAM_SOC_Raster1.tif")

plot(GCAM_SOC)

#Read in the 2017 Soil Carbon Data
SoilGrids_2017_100cm<-rast("OCSTHA_M_100cm_250m_ll.tif")

#Reproject into uniform raster format
terra::project(SoilGrids_2017_100cm, GCAM_SOC, method= "near", filename= "Reproj_2017_SOC_100cm.tif", overwrite=TRUE) 

SoilGrids_SOC<-rast("Reproj_2017_SOC_100cm.tif")

plot(SoilGrids_SOC)
```

## Differences

```{r}
difSOC<-SoilGrids_SOC - GCAM_SOC


cuts=c(-100,0,50,100,150,200,250,500) #set breaks
pal <- colorRampPalette(c("yellow","red"))

plot(difSOC, breaks=cuts, col = pal(7)) #plot with defined breaks

plot(Basin.ID, col=NA, add = TRUE)

plot(difSOC)


```


## Start Masking

You can also embed plots, for example:

```{r}
#Set up the temporary File for processing
terraOptions(tempdir="C:/Users/aliesch/OneDrive - Environmental Protection Agency (EPA)/Desktop/processing")

#read in raster masks
cropMask<-rast("cropMask.tif")

SOC2017_crops<- mask(SoilGrids_SOC, cropMask, 
                           inverse = FALSE, updatevalue = NA, filename = "SOC2017_crops.tif")

tmpFiles(current=TRUE, remove=TRUE)



forestMask<-rast("forestMask.tif")
SOC2017_forests<- mask(SoilGrids_SOC, forestMask, 
                           inverse = FALSE, updatevalue = NA, filename = "SOC2017_forest.tif")

tmpFiles(current=TRUE, remove=TRUE)


grassMask<-rast("grassMask.tif")
SOC2017_grass<- mask(SoilGrids_SOC, grassMask, 
                           inverse = FALSE, updatevalue = NA, filename = "SOC2017_grass.tif")

tmpFiles(current=TRUE, remove=TRUE)

pastureMask<-rast("pasture_mask.tif")
SOC2017_pasture<- mask(SoilGrids_SOC, pastureMask, 
                           inverse = FALSE, updatevalue = NA, filename = "SOC2017_pasture.tif")

tmpFiles(current=TRUE, remove=TRUE)

RIDMask<-rast("RID_Mask.tif")
SOC2017_RID<- mask(SoilGrids_SOC, RIDMask, 
                           inverse = FALSE, updatevalue = NA, filename = "SOC2017_RID.tif")

tmpFiles(current=TRUE, remove=TRUE)

shrubMask<-rast("shrubMask.tif")
SOC2017_shrub<- mask(SoilGrids_SOC, shrubMask, 
                           inverse = FALSE, updatevalue = NA, filename = "SOC2017_shrub.tif")

tmpFiles(current=TRUE, remove=TRUE)

tundraMask<-rast("tundra_mask.tif")
SOC2017_tundra<- mask(SoilGrids_SOC, tundraMask, 
                           inverse = FALSE, updatevalue = NA, filename = "SOC2017_tundra.tif")

tmpFiles(current=TRUE, remove=TRUE)

urbanMask<-rast("urbanMask.tif")
SOC2017_urban<- mask(SoilGrids_SOC, urbanMask, 
                           inverse = FALSE, updatevalue = NA, filename = "SOC2017_urban.tif")

tmpFiles(current=TRUE, remove=TRUE)


```

# Read in Basin Shapefiles
```{r}
#Read in the Region/Basin Shapefile
BasinShapeFile<-st_read("reg_basin_boundaries_moirai_landcells_3p1_0p5arcmin.shp") 

#Check the projection
st_crs(BasinShapeFile)

#reproject the Basin Shape File to align with the soil carbon raster
BasinShapeFile_igh<- st_transform(BasinShapeFile, CRS("+proj=igh +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs") )

#Use the dplyr package and the tidyverse notation to create the polygons to dissolve into. 

Basin.ID <- BasinShapeFile_igh %>% 
    group_by(basin_id) %>% 
    summarise() 

#check the basins overall

ggplot(Basin.ID) + geom_sf(aes(fill = 'Basins'))

```

#Use ExactExtract to calculate the median and Q3
```{r}



SOC2017_crops<-rast("SOC2017_crops.tif")

Basin.ID$cropMedian<- exact_extract(SOC2017_crops, Basin.ID, fun='quantile', quantiles= 0.5)


ggplot(Basin.ID) + geom_sf(aes(fill = cropMedian))

mapCrops<-vect(Basin.ID)

cropMask<-rast("cropMask.tif")


cropRaster<-rasterize(mapCrops, cropMask, field="cropMedian", background=NA)

SOC_2017_crops<- mask(cropRaster, cropMask, 
                           inverse = FALSE, updatevalue = NA, filename = "SOC2017_crops_median.tif", overwrite=TRUE)


```


```{r}

SOC2017_forest<-rast("SOC2017_forest.tif")

Basin.ID$forestMedian<- exact_extract(SOC2017_forest, Basin.ID, fun='quantile', quantiles= 0.5)


ggplot(Basin.ID) + geom_sf(aes(fill = forestMedian))

mapForest<-vect(Basin.ID)

forestMask<-rast("forestMask.tif")


forestRaster<-rasterize(mapForest, forestMask, field="forestMedian", background=NA)

SOC_2017_forest<- mask(forestRaster, forestMask, 
                           inverse = FALSE, updatevalue = NA, filename = "SOC2017_forest_median.tif", overwrite=TRUE)
```


```{r}
SOC2017_grass<-rast("SOC2017_grass.tif")

Basin.ID$grassMedian<- exact_extract(SOC2017_grass, Basin.ID, fun='quantile', quantiles= 0.5)


ggplot(Basin.ID) + geom_sf(aes(fill = grassMedian))

mapGrass<-vect(Basin.ID)

grassMask<-rast("grassMask.tif")


grassRaster<-rasterize(mapGrass, grassMask, field="grassMedian", background=NA)

SOC_2017_grass<- mask(grassRaster, grassMask, 
                           inverse = FALSE, updatevalue = NA, filename = "SOC2017_grass_median.tif", overwrite=TRUE)


```

```{r}
SOC2017_pasture<-rast("SOC2017_pasture.tif")

Basin.ID$pastureMedian<- exact_extract(SOC2017_pasture, Basin.ID, fun='quantile', quantiles= 0.5)


ggplot(Basin.ID) + geom_sf(aes(fill = pastureMedian))

mapPasture<-vect(Basin.ID)

pastureMask<-rast("pasture_mask.tif")


pastureRaster<-rasterize(mapPasture, pastureMask, field="pastureMedian", background=NA)

SOC_2017_pasture<- mask(pastureRaster, pastureMask, 
                           inverse = FALSE, updatevalue = NA, filename = "SOC2017_pasture_median.tif", overwrite=TRUE)
```

```{r}
SOC2017_RID<-rast("SOC2017_RID.tif")

Basin.ID$RIDMedian<- exact_extract(SOC2017_RID, Basin.ID, fun='quantile', quantiles= 0.5)


ggplot(Basin.ID) + geom_sf(aes(fill = RIDMedian))

mapRID<-vect(Basin.ID)

RIDMask<-rast("RID_Mask.tif")


RIDRaster<-rasterize(mapRID, RIDMask, field="RIDMedian", background=NA)

SOC_2017_RID<- mask(RIDRaster, RIDMask, 
                           inverse = FALSE, updatevalue = NA, filename = "SOC2017_RID_median.tif", overwrite=TRUE)
```

```{r}
SOC2017_Shrub<-rast("SOC2017_shrub.tif")

Basin.ID$shrubMedian<- exact_extract(SOC2017_Shrub, Basin.ID, fun='quantile', quantiles= 0.5)


ggplot(Basin.ID) + geom_sf(aes(fill = shrubMedian))

mapShrub<-vect(Basin.ID)

shrubMask<-rast("shrubMask.tif")


shrubRaster<-rasterize(mapShrub, shrubMask, field="shrubMedian", background=NA)

SOC_2017_shrub<- mask(shrubRaster, shrubMask, 
                           inverse = FALSE, updatevalue = NA, filename = "SOC2017_shrub_median.tif", overwrite=TRUE)
```


```{r}
SOC2017_Tundra<-rast("SOC2017_tundra.tif")

Basin.ID$tundraMedian<- exact_extract(SOC2017_Tundra, Basin.ID, fun='quantile', quantiles= 0.5)


ggplot(Basin.ID) + geom_sf(aes(fill = tundraMedian))

mapTundra<-vect(Basin.ID)

tundraMask<-rast("tundra_mask.tif")


tundraRaster<-rasterize(mapTundra, tundraMask, field="tundraMedian", background=NA)

SOC_2017_tundra<- mask(tundraRaster, tundraMask, 
                           inverse = FALSE, updatevalue = NA, filename = "SOC2017_tundra_median.tif", overwrite=TRUE)
```

```{r}
SOC2017_Urban<-rast("SOC2017_urban.tif")

Basin.ID$urbanMedian<- exact_extract(SOC2017_Urban, Basin.ID, fun='quantile', quantiles= 0.5)


ggplot(Basin.ID) + geom_sf(aes(fill = urbanMedian))

mapUrban<-vect(Basin.ID)

urbanMask<-rast("urbanMask.tif")


urbanRaster<-rasterize(mapUrban, urbanMask, field="urbanMedian", background=NA)

SOC_2017_urban<- mask(urbanRaster, urbanMask, 
                           inverse = FALSE, updatevalue = NA, filename = "SOC2017_urban_median.tif", overwrite=TRUE)
```


#Create a Polygon from the Median Value

#Rasterize Polygon

#Assemble the final Rasters



# Merge all Files, Phase 1

```{r}

grassTiff<-rast("SOC2017_grass_median.tif")
pastTiff<-rast("SOC2017_pasture_median.tif")

mosaic1<-mosaic(pastTiff, grassTiff, fun="mean", filename="mosaic1.tif", overwrite=TRUE, wopt=list())

tmpFiles(current=TRUE, remove=TRUE)

tundraTiff<-rast("SOC2017_tundra_median.tif")
urbanTiff<-rast("SOC2017_urban_median.tif")

mosaic2<-mosaic(tundraTiff, urbanTiff, fun="mean", filename="mosaic2.tif", overwrite=TRUE, wopt=list())

tmpFiles(current=TRUE, remove=TRUE)


forestTiff<-rast("SOC2017_forest_median.tif")
shrubTiff<-rast("SOC2017_shrub_median.tif")

mosaic3<-mosaic(forestTiff, shrubTiff, fun="mean", filename="mosaic3.tif", overwrite=TRUE, wopt=list())

tmpFiles(current=TRUE, remove=TRUE)

cropTiff<-rast("SOC2017_crops_median.tif")
RIDTiff<-rast("SOC2017_RID_median.tif")


mosaic4<-mosaic(cropTiff, RIDTiff, fun='mean', filename="mosaic4.tif", overwrite=TRUE, wopt=list())

tmpFiles(current=TRUE, remove=TRUE)
```

# Mosaic Files - Part 2

```{r}

mosaic1<-rast("mosaic1.tif")
mosaic2<-rast("mosaic2.tif")

semiMosaic1<-mosaic(mosaic1, mosaic2, fun='mean', filename="semimosaic1.tif", overwrite=TRUE, wopt=list())

tmpFiles(current=TRUE, remove=TRUE)



mosaic3<-rast("mosaic3.tif")
mosaic4<-rast("mosaic4.tif")

semiMosaic2<-mosaic(mosaic3, mosaic4, fun='mean', filename="semimosaic2.tif", overwrite=TRUE, wopt=list())

tmpFiles(current=TRUE, remove=TRUE)
```

### Compile the Final Pairing

```{r}
semiMosaic1<-rast("semimosaic1.tif")
semiMosaic2<-rast("semimosaic2.tif")

SoilGrids2017_Est<-mosaic(semiMosaic1, semiMosaic2, fun='mean', filename="SoilGrids2017_100cm_LU_Raster.tif", overwrite=TRUE, wopt=list())
```


#Comparisons
```{r}
GCAM<-rast("GCAM_SOC_Raster1.tif")
FAO<-rast("FAO_GSOCmap1.5.0.tif")
SoilGrids2017<-rast("SoilGrids2017_100cm_LU_Raster.tif")


terra::project(FAO, GCAM, method= "near", filename= "Reproject_FAO_100cm.tif")

FAO_Repro<-rast("Reproject_FAO_100cm.tif")

dif2SOC<-FAO_Repro-GCAM_SOC

cuts=c(-300,0,50,100,150,200,250,500) #set breaks

plot(dif2SOC, breaks=cuts, col = pal(7)) #plot with defined breaks
 
plot(Basin.ID, col=NA, add = TRUE)