---
title: "Land Carbon Recreation"
author: "Mandy Liesch"
date: "2023-02-08"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(terra)
library(tidyverse)
library(raster)
library(exactextractr)
library(sp)
library(rgdal)
library(sf)
library(tidyverse)

```



## Data Sources
### Data Files
GCAM_Basin_IDs.csv: This file has the basin names, and GLU, and basin number. 

GCAM 5.0 Core carbon Assumptions: This file has the managed and unmanaged soil carbon assumptions garnered from the land use in the [PNNL Technical Report](https://www.pnnl.gov/main/publications/external/technical_reports/PNNL-21025.pdf). The file has all different land uses, and will need to be manipulated and split by land use type. 

Moirai SoilGrids 2017 Q3 Estimates: This file was used for the new GCAM 6.0 update, and utilizes the soilgrids 2017 Third Quartile values to assume steady state functions.


### Raster Files
Land Use Raster: Created by Sarah Harrison, this is an in home modification of the European Space Agency/MODIS/Ramenkutty Raster. 

Values
1: Cropland
2: Forest
3: Grassland
4: Shrubland
5: Urban Land
6: RID
7: Tundra
8: Pasture


### Shapefiles
GCAM Basin Levels: There is an issue with the actual basin value level in the Moirai zip folder, so the region/boundary shapefile was used and dissolved to the Basin level in the igh projection. 

# Results


## Dissolve the Shapefile into Major Basins

```{r}
#Read in the Region/Basin Shapefile
BasinShapeFile<-st_read("reg_basin_boundaries_moirai_landcells_3p1_0p5arcmin.shp") 

#Check the projection
st_crs(BasinShapeFile)

#reproject the Basin Shape File to align with the soil carbon raster
BasinShapeFile_igh<- st_transform(BasinShapeFile, CRS("+proj=igh +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs") )

#Use the dplyr package and the tidyverse notation to create the polygons to dissolve into. 

Basin.ID <- BasinShapeFile_igh %>% 
    group_by(basin_id) %>% 
    summarise() 

#check the basins overall

ggplot(Basin.ID) + geom_sf(aes(fill = 'Basins'))
```

There are 232 different basins dissolved together. 

## Merge Basin Names to the New Shapefile

```{r}

basinNames<-read_csv("GCAM_Basin_IDs.csv")

basinNames$GLU_code<-NULL
basinNames$GLU_Name<-NULL

Basin.ID<-merge(basinNames, Basin.ID, by.x='Basin_ID', by.y='basin_id')

head(Basin.ID)
```

## Summarize the GCAM core Carbon Document by Land Use

```{r}
carbonC<-read_csv("GCAM 5.0 Core carbon assumptions.csv")

head(carbonC)

#remove the unnecessary columns 
carbonC$UniqueID<-NULL
carbonC$`Area ha`<-NULL

#Break the region and basin into two different columns
carbon<-separate(carbonC, RegionBasinID, c('Region', 'Basin'), sep='_', remove=TRUE)

#Remove the weird NA Columns
carbon<- carbon %>% drop_na()

head(carbon)
```


# Land Use Specific Processing

## Cropland

### Average together managed and unmanaged land by land type

```{r}

croplandC<- carbon %>% 
  filter(str_detect(carbon$GCAMLand, "Crop")) %>%
  group_by(Basin) %>%  
  summarise_at(vars(soilC), list(soilC = mean))

```


### Merge Datafile and Shapefile
```{r}
cropBasinID<-merge(croplandC, Basin.ID, by.x='Basin', by.y='Basin_ID')

maps<-st_as_sf(cropBasinID)

class(maps)

ggplot(maps) + geom_sf(aes(fill = soilC))

#convert to the raster
mapCrops<-vect(maps)

setwd("C:/Users/aliesch/OneDrive - Environmental Protection Agency (EPA)/Desktop/Intermediate Rasters/Historic Land Use/Recreations")

LandUse<-rast("Recreations/LU1700.tif")



```



### Use the land use raster to create a mask



```{r}
#Forest Classification
# classified as 2 in land use change


cropMask<- classify(LandUse, cbind(id=c(0, 1, 2, 3, 4, 5, 6, 7, 8, 26), 
                                      v=c(NA, 1, NA, NA, NA, NA, NA, NA, NA, NA)), filename='cropMask.tif')

plot(cropMask)

#convert to the raster
mapCrops<-vect(maps)

cropRaster<-rasterize(mapCrops, cropMask, field="soilC", background=NA)
```

### Extract Land Use Values into a New Raster

```{r}
SC_1700_crops<- mask(cropRaster, cropMask, 
                           inverse = FALSE, updatevalue = NA, filename = "SoilCRasters/SC_1700_crops.tif")
```

